---
title: "IGRAresults"
author: "Hannah Rickman"
date: "2023-02-08"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
.libPaths("C:/Users/HannahRickman/AppData/Local/R/win-library/4.2")
```

```{r}
library(tidyverse)
library(here)
library(lubridate)
library(table1)
library(epiR)
library(patchwork)
library(cowplot)
library(readr)
library(janitor)
library(rlang)
library(reshape)
library(viridis)
library(flextable)
library(openxlsx)
library(YesSiR)
library(broom)
library(arm)
library(cmdstanr)
set_cmdstan_path("C:/Users/HannahRickman/OneDrive - London School of Hygiene and Tropical Medicine/Documents/.cmdstan/cmdstan-2.34.1")
library(tidybayes)
library(brms)
library(broom.mixed)
library(RColorBrewer)
```


## 1. Reading & cleaning  

### Read in  

```{r}
igra <- read_csv("igra_cleaned.csv") 
```

### Cleaning - QFT-Plus results
```{r}
igra <- igra %>%
mutate(
  QFTP_result = factor(QFTP_result, levels = c("Negative", "Positive", "Indeterminate")),
  QFTP_TB2_minus_nil = QFTP_TB2 - QFTP_nil,
  QFTP_TB1_minus_nil = QFTP_TB1 - QFTP_nil,
  QFTP_mitogen_minus_nil = QFTP_mitogen - QFTP_nil,
  #set up as logical for epi tests
  qftp_bin = factor(case_when( 
    QFTP_result == "Positive" ~ 1, 
    QFTP_result == "Negative" ~0),
    levels=c(1,0), labels=c("QFTP+", "QFTP-")),
  #Define reasons for indeterminate results
  indet_reason = case_when(
    QFTP_result == "Indeterminate" & QFTP_nil <=8 & QFTP_mitogen_minus_nil <0.5 ~ "Low mitogen",
    QFTP_result == "Indeterminate" & QFTP_nil >8 ~ "High nil"),
#Define whether IGRAs are positive on TB1, TB2 or both
  tb1_pos = case_when(
    QFTP_TB1_minus_nil >= 0.35 & QFTP_TB1_minus_nil/QFTP_nil >=0.25 & QFTP_nil <=8.0 ~ "TB1 Pos",
          TRUE~"TB1 Neg"),
  tb2_pos = case_when(
    QFTP_TB2_minus_nil >= 0.35 & QFTP_TB2_minus_nil/QFTP_nil >=0.25 & QFTP_nil <=8.0 ~ "TB2 Pos",
          TRUE~"TB2 Neg"),
  tb1_tb2 = case_when(
    tb1_pos == "TB1 Pos" & tb2_pos == "TB2 Pos" ~ "TB1 & TB2 Positive",
    tb1_pos == "TB1 Pos" & tb2_pos == "TB2 Neg" ~ "TB1 only positive",
    tb1_pos == "TB1 Neg" & tb2_pos == "TB2 Pos" ~ "TB2 only positive",
    tb1_pos == "TB1 Neg" & tb2_pos == "TB2 Neg" ~ as.character(QFTP_result)), 
#Define whether positives are at >0.7
  igra_0.7 = case_when(
      QFTP_result == "Positive" & QFTP_TB1_minus_nil >= 0.7 ~ "IGRA positive at 0.7",
      QFTP_result == "Positive" & QFTP_TB2_minus_nil >= 0.7 ~ "IGRA positive at 0.7",
      QFTP_result == "Positive" & QFTP_TB1_minus_nil < 0.7 & QFTP_TB2_minus_nil <0.7 ~ "IGRA positive at 0.35-0.7",
      TRUE ~ as.character(QFTP_result)),
#Define whether negatives due to nil adjustment
#0.46 corresponds to 0.35 (the cut-off) plus the MEDIAN nil value
  nil_adjusted = case_when(
      QFTP_result == "Negative" & QFTP_TB2 >= 0.46 ~ "Negative after adjusted", 
      QFTP_result == "Negative" & QFTP_TB1 >= 0.46 ~ "Negative after adjusted",
      QFTP_result == "Negative" & QFTP_TB2 < 0.46 & QFTP_TB1 <0.46 ~ "Negative unadjusted"
      )) 
```


### Cleaning - QIAreach results
```{r}
igra <- igra %>% mutate(
  QIAreach_cat = as_factor(case_when(
    QIAreach_QFT == "Negative" ~ "Negative",
    QIAreach_QFT =="Positive" & QIAreach_TTP_s == 1200 ~ "Borderline positive, TTP 20 mins",
    QIAreach_QFT =="Positive" & QIAreach_TTP_s < 1200 ~ "Strong positive, TTP <20 mins")),
  QIAreach_poslevel = as_factor(case_when(
    QIAreach_QFT == "Negative" ~ NA_character_,
    QIAreach_QFT =="Positive" & QIAreach_TTP_s == 1200 ~ "Borderline positive, TTP 20 mins",
    QIAreach_QFT =="Positive" & QIAreach_TTP_s < 1200 ~ "Strong positive, TTP <20 mins")), 
  QIAreach_QFT = factor(QIAreach_QFT, levels = c("Negative", "Positive")),
#TTP
  QIAreach_TTP_min = QIAreach_TTP_s/60,
  ttp_inv_min = 1/QIAreach_TTP_min,
  ttp_inv_0 = case_when(
    QIAreach_QFT== "Negative" ~ 0,
    QIAreach_QFT== "Positive" & !is.na(QIAreach_TTP_s) ~ 1/QIAreach_TTP_s), 
#For epi tests set up as logical
  qia_01 = case_when(
    QIAreach_QFT == "Positive" ~ 1,
    QIAreach_QFT == "Negative" ~0),
  qia_bin= factor(qia_01, levels=c(1,0), labels=c("QIA+", "QIA-")),
    all="all") 
```

### Cleaning: Concordance

```{r}
igra <- igra %>% mutate(
  #Define overall result status, excluding indeterminates
  concordance_status =factor(case_when(
    QFTP_result == "Positive" & QIAreach_QFT == "Positive" ~ "True Positive",
    QFTP_result == "Positive" & QIAreach_QFT == "Negative" ~ "False Negative",
    QFTP_result == "Negative" & QIAreach_QFT == "Positive" ~ "False Positive",
    QFTP_result == "Negative" & QIAreach_QFT == "Negative" ~ "True Negative"), 
    levels = c("True Positive", "False Negative", "False Positive", "True Negative")),
  #Concordance status including indeterminates
  concordance_status_ind = factor(case_when(
    QFTP_result == "Positive" & QIAreach_QFT == "Positive" ~ "True Positive",
    QFTP_result == "Positive" & QIAreach_QFT == "Negative" ~ "False Negative",
    QFTP_result == "Negative" & QIAreach_QFT == "Positive" ~ "False Positive",
    QFTP_result == "Negative" & QIAreach_QFT == "Negative" ~ "True Negative",
    QFTP_result == "Indeterminate" & QIAreach_QFT == "Negative" ~ "Indeterminate QFT Plus, Negative QIAreach",
    QFTP_result == "Indeterminate" & QIAreach_QFT == "Positive" ~ "Indeterminate QFT Plus, Positive QIAreach"),
    levels = c("True Positive", "False Negative", "False Positive", "True Negative", "Indeterminate QFT Plus, Positive QIAreach", "Indeterminate QFT Plus, Negative QIAreach"))
)
```

### Cleaning: CoVs
```{r}
igra <- igra %>% mutate(
#Age categories - here using 19 as relevant cut-off; note there are no participants aged 5-10
  age_cat = cut(age, breaks = c(0, 6, 19.999, 41), 
                       labels = c("1-4 years", "10-19 years", "20-40 years"), 
                                  closed=LEFT),
  tbrx_bin = case_when(tbrx_cat == "Unsure" ~ "No previous TB treatment",
                              tbrx_cat == "No" ~ "No previous TB treatment",
                              tbrx_cat == "Yes" ~ "Previous TB treamtent"),
  bcg_bin = case_when(bcg_ind == "Unsure" ~ "No BCG vaccination",
                             bcg_ind == "No" ~ "No BCG vaccination",
                             bcg_ind == "Yes" ~ "Previous BCG vaccination"), 
  currentcough_ind = case_when(currentcough_ind == "Yes" ~ "Current cough",
                                   currentcough_ind == "No" ~ "No current cough"),
  any_tbsx = case_when(any_tbsx == "Yes" ~ "Current TB symptoms",
                            any_tbsx == "No" ~ "No current TB symptoms"),
  muac_cat = case_when(muac <12.5 ~ "<12.5",
                                muac>=12.5 & muac<13.5 ~ "12.5 to <13.5",
                              muac>=13.5 ~ ">=13.5"),    
  muac_bin = case_when(muac <13.5 ~ "Low",
                              muac>=13.5 ~ "Normal"), 
  hivexp_bin = case_when(hiv_exposed == "HIV Unexposed" ~ "HIV Unexposed",
                                hiv_exposed == "HIV Exposed Uninfected" ~ "HIV Exposed",
                                hiv_exposed == "HIV Positive" ~ "HIV Exposed"))


label(igra$age) <- "Age (years) - median (IQR)"
label(igra$sex) <- "Sex" 
label(igra$rec_location) <- "Recruitment location"
label(igra$hivstatus_binary) <- "HIV status"
label(igra$hiv_exposed) <- "HIV exposure"
label(igra$currentcough_ind) <- "Current cough"
label(igra$any_tbsx) <- "Current TB symptoms"
label(igra$bcg_bin) <- "BCG vaccination status"
label(igra$tbrx_bin) <- "Previous TB treatment"
label(igra$QFTP_result) <- "QFT-Plus result"
label(igra$QIAreach_QFT) <- "QIAreach QFT result"
label(igra$QIAreach_poslevel) <- "Positive QIAreach TTP*"
label(igra$muac_cat) <- "MUAC"

```


## 2. Descriptive stats  - Table 1  
```{r}
table1<- table1(~age + sex + rec_location + 
       hivstatus_binary + hiv_exposed + 
         currentcough_ind + any_tbsx + bcg_bin + muac_cat +
         tbrx_bin + 
         QFTP_result + QIAreach_QFT + QIAreach_poslevel
         | age_cat, data=igra, 
        render.continuous=c(.="Median (Q1-Q3)"), 
       render.missing = NULL,
       render.categorical = "FREQ (PCTnoNA%)",
       footnote="* Data missing for one participant. IQR: Interquartile range. TB: tuberculosis. TTP: time to positivity")
table1 #To do: tidy this
```


## 3. Table 2 - concordance, sens/spec/PPV/NPV

### Contingency table  
```{r}
igra %>% 
  group_by(qia_bin, QFTP_result) %>% 
  summarise(n=n()) %>%
  pivot_wider(id_cols = QFTP_result, names_from = qia_bin, values_from = n)
```
  
### First part of table 2    
```{r}
conttable_all <- igra %>% 
  tabyl(concordance_status_ind) %>% dplyr::select(-percent) %>% dplyr::rename(All = n)
conttable_age <- igra %>%
  tabyl(concordance_status_ind, age_cat)
conttable_hiv <- igra %>% filter(age>=20) %>% 
    tabyl(concordance_status_ind, hivstatus_binary)


conttable<- conttable_all %>% left_join(conttable_age, by="concordance_status_ind") %>%
  left_join(conttable_hiv, by="concordance_status_ind")
# conttable
```

### Function for caclulating sensitivty, specificity, PPV/NPV and kappa across strata  

```{r}
acc_func <- function(data, truth, test, strata) {
  
  strata_sym <- ensym(strata)
  strata_name <- as_name(strata_sym)  # Convert strata_sym to character name
  
  big_tab <- data %>%
    group_by(!!strata_sym) %>%
    nest() %>% #Creates a nested dataframe within a dataframe - for each strata group - which lives inside the new "data" column
    mutate(
      cross_tab = map(data, ~ table(.x[[{{test}}]], .x[[{{truth}}]])), #Create a new variable cross_tab
                    #map() function iterates over each element in the "data" column of the nested dataframe
                    ##For each individual nested dataframe within data colmn (.x) this creates a contingency table - extracts values of test columns from nested df and generates a table of counts for combinations of those values
      chi_sq = map(cross_tab, ~ chisq.test(.)), #Across the cross_tab column, which contains the contingency tables, perform a chi squared test                       
      p_value = map_dbl(chi_sq, ~ ..1[["p.value"]]),
      metrics = map(cross_tab, ~ epi.tests(.)), #Perform epi tests across each element in the cross_tab column
      kappa = map(cross_tab, ~epi.kappa(.)), #
      n = map_dbl(data, nrow)) %>%
    mutate(strata_level = !!strata_sym)
  
  kappa_df <- map2(big_tab$kappa, unique(data[[as_string(strata_sym)]]), ~ {
    tibble(
      strata = as_string(strata_sym),
      strata_level = .y,
      .x$kappa
    )
  }) %>%
    map_df(identity) %>% dplyr::select(-se) %>% mutate(statistic = "kappa")
  
  epi_df <- map2(big_tab$metrics, unique(data[[as_string(strata_sym)]]), ~ {
    tibble(
      strata = as_string(strata_sym),
      strata_level = .y,
      .x$detail
    )
  }) 
  
  results_df <- kappa_df %>% bind_rows(epi_df) %>%
  pivot_wider(id_cols = c("strata", "strata_level"), names_from="statistic", values_from=c("est", "upper", "lower")) %>%
    dplyr::select(strata, strata_level, ends_with("_se"), ends_with("_sp"), ends_with("pv.pos"), ends_with("pv.neg"), ends_with("kappa")) 

final_tab <- big_tab %>% dplyr::select(strata_level, n) %>% left_join(results_df, by="strata_level") %>% #Reformat for table
  mutate(across(
    ends_with(c("se", "sp", "pv.pos", "pv.neg")),
    ~scales::percent(.x, scale = 100, accuracy=1))) %>%
  mutate(across(ends_with("kappa"), ~format(round(., digits=2), nsmall = 2))) %>%
   mutate(Sensitivity = glue::glue("{est_se} ({lower_se}-{upper_se})"),
          Specificity = glue::glue("{est_sp} ({lower_sp}-{upper_sp})"),
          PPV = glue::glue("{est_pv.pos} ({lower_pv.pos}-{upper_pv.pos})"),
          NPV = glue::glue("{est_pv.neg} ({lower_pv.neg}-{upper_pv.neg})"),
          Kappa = glue::glue("{est_kappa} ({lower_kappa}-{upper_kappa})")
          ) %>%
  dplyr::select(strata, strata_level, n, Sensitivity, Specificity, PPV, NPV, Kappa)

return(final_tab)   


}
```

### 2nd part of table 2
```{r}
#acc_func(data=igra, truth="qftp_bin", test="qia_bin", strata = age_cat) 
```

### Table 2  
Putting together results for several different strata into one table
```{r}
subgroup_acc <- acc_func(data=igra, truth="qftp_bin", test="qia_bin", strata = all) %>% bind_rows(
acc_func(data=igra, truth="qftp_bin", test="qia_bin", strata = age_cat)) %>% bind_rows(
igra %>% filter(age>=20) %>% acc_func(truth="qftp_bin", test="qia_bin", strata = hivstatus_binary)) %>% ungroup() %>%
  dplyr::select(strata, strata_level, n, Sensitivity, Specificity, PPV, NPV, Kappa) 
subgroup_acc
```

  
Transpose for Table  
```{r}
tab2_acc<- subgroup_acc %>% t() %>% as.data.frame() %>% row_to_names(2) %>% tibble::rownames_to_column() %>% 
  dplyr::rename(All = all)
tab2_acc

#To get table into right format: could actually move this all into the function above
table2<- conttable %>% as.data.frame() %>% dplyr::rename(rowname = concordance_status_ind) %>%
  mutate(across(where(is.numeric), as.character)) %>% bind_rows(tab2_acc) %>% 
  mutate(row_number = case_when(rowname == "n"~ 0)) %>% arrange(row_number) %>% dplyr::select(-row_number)
knitr::kable(table2, format = "markdown")
```

### Other CoVs - supplementary data  


```{r}
conttable_hivexp <- igra %>% filter(age<=5) %>%
  tabyl(concordance_status_ind, hivexp_bin)
conttable_muac <- igra %>% filter(age<=5) %>%
  tabyl(concordance_status_ind, muac_bin)
conttable_all_u5 <- igra %>% filter(age<=5) %>%
  tabyl(concordance_status_ind) %>% dplyr::select(-percent) %>% dplyr::rename(All = n)


conttable_supp<- conttable_all_u5 %>% left_join(conttable_muac, by="concordance_status_ind") %>%
  left_join(conttable_hivexp, by="concordance_status_ind")

#Putting together results for several different strata into one table
subgroup_acc_supp <- (
igra %>% filter(age<=5) %>% acc_func(truth="qftp_bin", test="qia_bin", strata = all)) %>% 
  bind_rows(
igra %>% filter(age<=5) %>% acc_func(truth="qftp_bin", test="qia_bin", strata = muac_bin)) %>%
  bind_rows(
igra %>% filter(age<=5) %>% acc_func(truth="qftp_bin", test="qia_bin", strata = hivexp_bin)) %>% ungroup() %>%
  dplyr::select(strata, strata_level, n, Sensitivity, Specificity, PPV, NPV, Kappa) 
subgroup_acc_supp


#Transpose for Table
tab2_acc_supp<- subgroup_acc_supp %>% t() %>% as.data.frame() %>% row_to_names(2) %>% tibble::rownames_to_column() %>% 
  dplyr::rename(All = all)
tab2_acc_supp

conttable_supp %>% as.data.frame() %>% dplyr::rename(rowname = concordance_status_ind) %>%
  mutate(across(where(is.numeric), as.character)) %>% bind_rows(tab2_acc_supp) %>% 
  mutate(row_number = case_when(rowname == "n"~ 0)) %>% arrange(row_number) %>% dplyr::select(-row_number)  #%>% write.xlsx("supp_table_2.xlsx")
conttable_supp
```


## 4. Logistic regression for sensitivity and specificity    

```{r}
#Define QFT Pos and Neg subgroups (as sensitivity defined in those with positive QFT-Plus, and specificity is defined in those with negative QFT-Plus )
qftp_pos <- igra %>% filter(qftp_bin == "QFTP+")
qftp_neg <- igra %>% filter(qftp_bin == "QFTP-")
```

### i. By age category     
#### A. Sensitivity

Sensitivity = P(positive QIAreach | positive QFT-Plus)  

How does sensitivity relate to age?    
```{r}
ggplot(qftp_pos, aes(age, as.numeric(qia_01))) +
  geom_point() +
  geom_smooth(span=1) +
    theme_bw() +
  ylim(0,1) +
  scale_y_continuous(name="Sensitivity", breaks=seq(0,1,0.25))+
  labs(x = "Age, years")
```
  
Linear is visually reasonable (clearly will fail at extremes)  


```{r}
sens_glm_age_cat <- glm(qia_01 ~ age_cat, data = qftp_pos, family = binomial)
summary(sens_glm_age_cat)
```
What if age instead modelled continously?   
```{r}
sens_glm_age_cont <- glm(qia_01 ~ age, data = qftp_pos, family = binomial)
summary(sens_glm_age_cont)
```
AIC is actually lower with continuous age.  
However for descriptive purposes for manuscript (particularly as there is a "missing" age category) it may be most logical to categorise - especially if relationship with specificity is not as clear. 

```{r}
exp(cbind(OR = coef(sens_glm_age_cont), confint(sens_glm_age_cont)))
```

#### B. Specificity  
How does specificity relate to age?
```{r}
ggplot(qftp_neg, aes(age, 1-as.numeric(qia_01))) +
  geom_point() +
  geom_smooth(span=1) +
  theme_bw() +
  labs(x = "Age, years",
       y = "Specificity")
```

Less strongly correlated; linear not as good. 
Note flipped axes - we are seeing specificity DECREASE with age


```{r}
spec_glm_age_cat <- glm(qia_01 ~ age_cat, data = qftp_neg, family = binomial)
summary(spec_glm_age_cat)
```
```{r}
spec_glm_age_cont <- glm(qia_01 ~ age, data = qftp_neg, family = binomial)
summary(spec_glm_age_cont)
```
Marginally lower AIC with continuous age, but again close, and appears less linear visually

### ii. By other CoVs  
- Some other CoVs only tested in certain age groups
- E.g. MUAC / HIV exposure / BCG in U5s
- HIV very uncommon in U20s  


Define QFT Pos and Neg subgroups by age
```{r}
qftp_pos_u5 <- igra %>% filter(qftp_bin == "QFTP+") %>% filter(age<=5)
qftp_pos_o19 <- igra %>% filter(qftp_bin == "QFTP+") %>% filter(age>=20)
qftp_neg_u5 <- igra %>% filter(qftp_bin == "QFTP-") %>% filter(age<=5)
qftp_neg_o19 <- igra %>% filter(qftp_bin == "QFTP-") %>% filter(age>=20)
```


#### A. Sensitivity and specificity by HIV status

```{r}
sens_glm_hiv <- glm(qia_01 ~ hivstatus_binary, data = qftp_pos_o19, family = binomial)
summary(sens_glm_hiv)
```

```{r}
spec_glm_hiv <- glm(qia_01 ~ hivstatus_binary, data = qftp_neg_o19, family = binomial)
summary(spec_glm_hiv)
```

#### B. Sensitivity and specificity by HIV exposure status
```{r}
sens_glm_hivexp <- glm(qia_01 ~ hivexp_bin, data = qftp_pos_u5, family = binomial)
summary(sens_glm_hivexp)

spec_glm_hivexp <- glm(qia_01 ~ hivexp_bin, data = qftp_neg_u5, family = binomial)
summary(spec_glm_hivexp)
```

  
#### C. Sensitivity and specificity by MUAC  

How does sensitivity relate to MUAC?  
```{r}
ggplot(qftp_pos_u5, aes(muac, as.numeric(qia_01))) +
  geom_point() +
  geom_smooth(span=1)
```

```{r}
ggplot(qftp_neg_u5, aes(muac, -as.numeric(qia_01))) +
  geom_point() +
  geom_smooth(span=1)
```

No clear relationship, however given MUAC conventionally categorised:  
```{r}
sens_glm_muac <- glm(qia_01 ~ muac_bin, data = qftp_pos_u5, family = binomial)
summary(sens_glm_muac)

spec_glm_muac <- glm(qia_01 ~ muac_bin, data = qftp_neg_u5, family = binomial)
summary(spec_glm_muac)
```

## 5. Possible sources of misclassification  

### A. Why do indeterminates occur?  

```{r}
igra %>% filter(QFTP_result == "Indeterminate") %>% tabyl( indet_reason, age_cat) %>% adorn_totals("col")  %>% adorn_percentages("col") %>% 
  adorn_pct_formatting(rounding = "half up", digits = 0) 
```

Mostly due to low mitogen, especially in the young kids  

```{r}
igra %>% filter(QFTP_result == "Indeterminate") %>% tabyl(QIAreach_QFT, age_cat, indet_reason)%>% adorn_totals("col") %>% adorn_percentages("col") %>% 
  adorn_pct_formatting(rounding = "half up", digits = 0) 

```

### B. What happens if indeterminates are reclassified?  
#### i. To positive
```{r}
igra <- igra %>%
  mutate(qftp_reclassify_indpos = case_when(
    QFTP_result == "Negative" ~ 0,
    QFTP_result == "Positive" ~ 1,
    QFTP_result == "Indeterminate" ~ 1), 
        qftp_reclassify_indpos = factor(qftp_reclassify_indpos, levels=c(1,0), labels = c("Pos or Ind", "Neg")))
    
reclassify_indpos_acc <- acc_func(data=igra, truth="qftp_reclassify_indpos", test="qia_01", strata = all) %>% 
  bind_rows(acc_func(data=igra, truth="qftp_reclassify_indpos", test="qia_01", strata = age_cat)) %>% 
  bind_rows(igra %>% filter(age>=20) %>% acc_func(truth="qftp_reclassify_indpos", test="qia_01", strata = hivstatus_binary))  %>% 
    t() %>% as.data.frame() %>% row_to_names(3) %>% tibble::rownames_to_column() %>% 
  dplyr::rename(All = all)
```
  
#### ii. To negative
```{r}
igra <- igra %>%
  mutate(qftp_reclassify_indneg = case_when(
    QFTP_result == "Negative" ~ 0,
    QFTP_result == "Positive" ~ 1,
    QFTP_result == "Indeterminate" ~ 0), 
        qftp_reclassify_indneg = factor(qftp_reclassify_indneg, levels=c(1,0), labels = c("Pos or Ind", "Neg")))
    
reclassify_indneg_acc <- acc_func(data=igra, truth="qftp_reclassify_indneg", test="qia_01", strata = all) %>% 
  bind_rows(acc_func(data=igra, truth="qftp_reclassify_indneg", test="qia_01", strata = age_cat)) %>% 
  bind_rows(igra %>% filter(age>=20) %>% acc_func(truth="qftp_reclassify_indneg", test="qia_01", strata = hivstatus_binary))  %>% 
    t() %>% as.data.frame() %>% row_to_names(3) %>% tibble::rownames_to_column() %>% 
  dplyr::rename(All = all)
reclassify_indneg_acc 
```




### B. Results which are positive only on TB1
```{r}
igra %>% filter(QFTP_result == "Positive") %>% tabyl(tb1_tb2, age_cat) %>% 
  adorn_totals("col")%>% 
  adorn_percentages("col") %>%
    adorn_pct_formatting() %>%
  adorn_ns(position = "front") 
```

- Many of the young kids have positive results only on TB1 or TB2, not both
- Lower percentage in older age groups  

Of those which are positive only on TB1, what is the QIAreach result?  
```{r}
igra %>% filter(QFTP_result == "Positive") %>% 
  filter(tb1_tb2=="TB1 only positive") %>%
  tabyl(QIAreach_QFT,age_cat)%>% 
  adorn_totals("col")%>% 
  adorn_percentages("col") %>%
    adorn_pct_formatting() %>%
  adorn_ns(position = "front")
```

So of 59 "false negative" results (positive QFT-Plus, negative QIAreach), potentially 18/59 (30.5%) explained by a TB1-only positive result. 


### C. Low level positives on QFT-Plus
```{r}
igra %>% filter(QFTP_result == "Positive") %>% tabyl(QIAreach_QFT, igra_0.7, age_cat) 
```

### D. QFT-Plus esults which have a TB1/TB2 above the positive threshold, but then are once nil subtracted
```{r}
igra %>% filter(QFTP_result == "Negative") %>% tabyl(nil_adjusted, age_cat) %>% 
  adorn_totals("col")%>% 
  adorn_percentages("col") %>%
    adorn_pct_formatting() %>%
  adorn_ns(position = "front")  
```


```{r}
igra %>% filter(QFTP_result == "Negative") %>% 
  filter(nil_adjusted=="Negative after adjusted") %>%
  tabyl(QIAreach_QFT,age_cat)%>% 
  adorn_totals("col")%>% 
  adorn_percentages("col") %>%
    adorn_pct_formatting() %>%
  adorn_ns(position = "front") 

```
49 results with a high TB1 or TB2 but which when adjusted would be below the cut-off (6.1% of total) - most of these (32, 65%) are called as positive by QIAreach

## 6. Exploring quantitative QIAreach values as sources of misclassification  

```{r}
igra %>% 
  filter(QFTP_result !="Indeterminate") %>% tabyl(QIAreach_cat, QFTP_result) %>% 
 adorn_totals("col")%>% 
  adorn_percentages("row") %>%
    adorn_pct_formatting() %>%
  adorn_ns(position = "front") # %>% View() #write.xlsx("SuppTab_ttp20.xlsx")
```
74/149 "strong positive" QIAreach have a positive QFTP = 50%
23/156 "weak positive" QIAreach have a positive QFTP = 15%
59/647 "negative" QIAreach have a posiive QFTP = 9%

```{r}
igra %>% filter(concordance_status == "False Positive") %>%
  tabyl(QIAreach_cat)

```

Reclassifying "weak positives" as negative
```{r}
igra <- igra %>%
  mutate(qia_reclassify = case_when(QIAreach_cat == "Strong positive, TTP <20 mins" ~ 1,
                                    QIAreach_cat == "Negative" ~ 0,
                                    QIAreach_cat == "Borderline positive, TTP 20 mins" ~ 0),
  qia_reclassify = factor(qia_reclassify, levels=c(1,0), labels=c("Strong+", "Weak+or-")))  

reclassify_acc <- acc_func(data=igra, truth="qftp_bin", test="qia_reclassify", strata = all) %>% 
  bind_rows(acc_func(data=igra, truth="qftp_bin", test="qia_reclassify", strata = age_cat)) %>% 
  bind_rows(igra %>% filter(age>=20) %>% acc_func(truth="qftp_bin", test="qia_reclassify", strata = hivstatus_binary))  %>% 
    t() %>% as.data.frame() %>% row_to_names(3) %>% tibble::rownames_to_column() %>% 
  dplyr::rename(All = all)
```
  
Exploring QFTP values for each
```{r}
igra %>%  filter(QIAreach_cat != "NA_") %>% tabyl(igra_0.7, QIAreach_cat) %>% adorn_totals("col")
```
```{r}
igra %>% tabyl(igra_0.7, QIAreach_QFT) %>% adorn_totals("col")
```
84/123 "strong" positive IGRAs are positive by QIAreach = 68%
13/33 "weak" positive IGRAs are positive by QIAreach = 39%
209/797 negative IGRAs are positive by QIAreach = 26%


## 7. Comparing actual quantitative values

Removing indeterminate values
```{r}
igra_noind <- igra %>% filter(QFTP_result !="Indeterminate")
```

```{r}
unadjusted_tb2_graph_crude<- 
  igra_noind %>%
  filter(QIAreach_QFT=="Positive") %>%
ggplot(aes(x = QFTP_TB2, y = QIAreach_TTP_min)) +
  geom_point(alpha=0.7,
             aes(col=QFTP_result)) + 
  theme_bw() +
  theme(legend.position = "nil" ) + 
  facet_grid(.~age_cat) +
  labs(y=bquote("QIAreach: TTP(min)"),
         x="",
        title = "a",
        # x="QFT-Plus: Unadjusted TB2 (IU/mL)",
       col="QFT-Plus result") +
  ylim(0,20)
unadjusted_tb2_graph_crude<- unadjusted_tb2_graph_crude + 
  geom_smooth(span=1, linewidth = 0.75, colour = "slategray", alpha = 0.3)
```

Alternatively, plotting 1/TTP rather than TTP  
- a priori, we might expect the CONCENTRATION of IFN-g to be correlated with the RATE of QIAReach becoming positive (i.e. 1/TTP)
```{r}
unadjusted_tb2_graph_inv<- 
  igra_noind %>%
  filter(QIAreach_QFT=="Positive") %>%
ggplot(aes(x = QFTP_TB2, y = ttp_inv_min)) +
  geom_point(alpha=0.7,
             aes(col=QFTP_result)) + 
  theme_bw() +
  theme(legend.position = "bottom") + 
  facet_grid(.~age_cat) +
  labs(y=bquote("QIAreach: 1/TTP(" ~min^-1~")"),
         x="QFT-Plus: Unadjusted TB2 (IU/mL)",
       col="QFT-Plus result",
        title = "b")  
unadjusted_tb2_graph_inv<-unadjusted_tb2_graph_inv + geom_smooth(span=1, colour = "slategray", linewidth = 0.75, alpha = 0.3) 
```
```{r}
unadjusted_tb2_graph_crude /
  unadjusted_tb2_graph_inv
```

```{r}  
ggsave("Fig1_dual.png")
```
  
Neither of these appear linear in their relationship. Overall data (no age categorisation) for completeness.

```{r}
unadjusted_tb2_graph_inv_all <- igra_noind %>%
  filter(QIAreach_QFT=="Positive") %>%
ggplot(aes(x = QFTP_TB2, y = ttp_inv_min)) +
  geom_point(alpha=0.7,
             aes(col=QFTP_result)) + 
  theme_bw() +
  theme(legend.position = "bottom") + 
  labs(y=bquote("QIAreach: 1/TTP(" ~min^-1~")"),
         x="QFT-Plus: Unadjusted TB2 (IU/mL)",
       col="QFT-Plus result") 
unadjusted_tb2_graph_inv_all + 
  geom_smooth(span=1, colour = "slategray", linewidth = 0.75, alpha = 0.3)
```

## 8. Measurement of correlation  

- Aim to quantify the degree of correlation, by age-group
- Linear regression / R-squared not appropriate, as definitely non-linear
- We can therefore use a Spearman rank coefficient instead (requires only monotonic relationship)
- Note that transforming - i.e. taking 1/TTP - doesn't make any difference to the outcome of the statistical test as it's a ranked test  



```{r}
corr_all <-igra_noind %>%
  filter(QIAreach_QFT == "Positive")  
corr_u5 <- igra_noind %>%
  filter(QIAreach_QFT == "Positive") %>% filter(age_cat == "1-4 years")
corr_10 <- igra_noind %>%
  filter(QIAreach_QFT == "Positive") %>% filter(age_cat == "10-19 years")
corr_20 <- igra_noind %>%
  filter(QIAreach_QFT == "Positive") %>% filter(age_cat == "20-40 years")


cor.test(corr_u5$QFTP_TB2, corr_u5$QIAreach_TTP_s, method = "spearman") %>% tidy() %>% mutate(age_cat = "1-5 years") %>%
  bind_rows(
cor.test(corr_10$QFTP_TB2, corr_10$QIAreach_TTP_s, method = "spearman") %>% tidy() %>% mutate(age_cat = "10-19 years")) %>% 
  bind_rows(
cor.test(corr_20$QFTP_TB2, corr_20$QIAreach_TTP_s, method = "spearman") %>% tidy() %>% mutate(age_cat = "20-40 years")) %>%   
mutate(rho = estimate) %>% dplyr::select(age_cat, rho, statistic, p.value) 

```



## 9. Hurdle categorical model

Putting this all into one model
- Incorporating negatives in a hurdle distribution
```{r}
igra_noind <- igra_noind %>%
  mutate(
    ttp_cat2 = 2*floor(QIAreach_TTP_s/120)) %>% #Divide ttp into 2-minute brackets
  mutate(  
  ttp_cat2 = factor(
    case_when(QIAreach_QFT == "Negative"~ 0, #Create hurdle
                                   TRUE ~ ttp_cat2), 
              ordered = TRUE, 
              labels = c("Negative", "2 to <4", "4 to <6", "6 to <8", "8 to <10", 
                                      "10 to <12", "12 to <14", "14 to <16", "16 to <18", "18 to <20", "20")))

tabyl(igra_noind$ttp_cat2)
# 1 positive QIAreach is missing a TTP and theefore excluded
```

```{r}
igra_noind %>%
  ggplot() +
  geom_bar(aes(x=ttp_cat2))
```
Note that 0 here is negative - whereas it's actually logically "after" 20. 
But this doesn't matter for purposes of this model. 
We later designate 0 as our hurdle - and clearing the hurdle is an independent process to the gamma part


```{r}
set.seed(279)
m1 <- brm(
  data = igra_noind,
  family = hurdle_cumulative(link="cauchit"),
  bf(ttp_cat2 ~ QFTP_TB2*age_cat,
  hu ~ QFTP_TB2*age_cat),
  cores = 4, 
  chains = 4)
```

```{r}
plot(m1)
```

```{r}
summary(m1)
```

```{r}
pp_check(m1)
```

```{r}
pp_check(m1, type="bars")
```



```{r}
nd <- crossing(age_cat = unique(igra_noind$age_cat),
               QFTP_TB2 = modelr::seq_range(igra_noind$QFTP_TB2, n=20))


m1_out <- nd %>%
  add_epred_draws(m1, allow_new_levels=TRUE)

overall_outs_m1 <- m1_out %>%
  # ungroup() %>%
  # group_by(age_cat, QFTP_TB2) %>%
  mean_qi(.epred) %>%
  mutate(.category = fct_rev(fct_relevel(.category, "Negative", after=Inf)))

levels(overall_outs_m1$.category)

nb_cols2 <- 13
myblues2 <- colorRampPalette(brewer.pal(9, "Blues"))(nb_cols2)[4:13]
myblues2_rev <-rev(myblues2)
mycolors2 <- append(myblues2, "#C60C30",after = 0)

m1_plot <- overall_outs_m1 %>%
  ggplot() +
  geom_area(aes(x=QFTP_TB2, y=.epred, fill=factor(.category)), size=.25, colour="white", linetype=2) +
  facet_grid(.~age_cat) +
  theme_bw() +
  scale_fill_manual(values = mycolors2,
                    name = "QIAreach TTP (minutes)") +
  scale_y_continuous(labels = scales::percent) +
  theme(legend.position = "bottom") + 
    guides(fill = guide_legend(nrow = 3)) +
  labs(y= "% QIAreach results within TTP interval", 
       x="QFT-Plus: Unadjusted TB2 (IU/mL)")

m1_plot
```

Improving the legend 
```{r}
legend_scale <- cbind(mycolors2, levels(overall_outs_m1$.category)) %>%
  as.data.frame()
legend_scale <- legend_scale[-1,] %>%
  mutate(val = as.numeric(substr(V2, 1,2))+1)

midpoints_list <- as.list(legend_scale$val)
colours_list <- as.list(legend_scale$mycolors2)

# Plot for the color gradient legend
legend_plot <- ggplot(legend_scale, aes(x = 1, y = val, fill = val)) +
  geom_tile() +
  scale_fill_gradientn(colours = rev(colours_list), guide = "colorbar",
                         limits = c(2,20), breaks = c(5, 10, 15, 20)) +
  guides(fill = guide_colorbar(direction="horizontal",
                               reverse=TRUE)) +
    theme(legend.text = element_text(size = 10),
          legend.title = element_text(size = 12)) +
  labs(fill = "TTP of positive QIAreach results (min)")

# Print the legend plot
legend_scale<- get_legend(legend_plot)
# plot(legend_scale)
```

```{r}
legend_plot2<- igra_noind %>%
  mutate(qia_res= factor(qia_01, levels=c(0,1), labels=c("Negative", "Positive"))) %>%
  ggplot() +
  geom_bar(aes(x= qia_res, fill=qia_res)) + 
  scale_fill_manual(values=c("#C60C30", "#4F9BCB")) +
  labs(fill = "QIAreach result") +
    theme(legend.direction="horizontal", 
          legend.text = element_text(size = 10),
          legend.title = element_text(size = 12))
legend_fill<- get_legend(legend_plot2)
#plot(legend_fill)

```


Put it all together

```{r}
#1. Plot without legend
m1_plot_nolegend <- overall_outs_m1 %>%
  ggplot() +
  geom_area(aes(x=QFTP_TB2, y=.epred, fill=factor(.category)), size=.25, colour="white", linetype=2) +
  facet_grid(.~age_cat) +
  theme_bw() +
  scale_fill_manual(values = mycolors2,
                    name = "QIAreach TTP (minutes)") +
  scale_y_continuous(labels = scales::percent) +
  theme(legend.position = "none") +
  labs(y= "% QIAreach results within TTP interval", 
       x="QFT-Plus: Unadjusted TB2 (IU/mL)")

legend_comb<- plot_grid(legend_fill, legend_scale, ncol=2)
#legend_comb
```

```{r}
comb_plot <- plot_grid(m1_plot_nolegend,legend_comb, ncol=1,
                       rel_heights = c(1, .2))
comb_plot
```

```{r}
ggsave("Fig2.png", width = 10, bg="white")
```


Pulling out some illustrative numbers  
```{r}
nd_test <- crossing(age_cat = unique(igra_noind$age_cat),
               QFTP_TB2 = c(0,1,5))
nd_test
m1_out_test <- nd_test %>%
  add_epred_draws(m1, allow_new_levels=TRUE) 
m1_out_test %>%
  mean_qi()

```

Recategorising output of same model (m1) to just include negative, positive at <20 mins and positive at 20 mins, to get 95% UIs

```{r}
m1_out_test %>% 
  mutate(.category = case_when(
    .category == "Negative" ~ "Negative",
    .category == "20" ~ "20 mins", #recategorise TTP
    TRUE ~ "<20 mins")) %>%  
  group_by(.row, age_cat, QFTP_TB2, .category) %>%
  summarise( #Manually performing function of mean_qi()
    .epred_mean = mean(.epred),                   # Mean prediction
    .epred_low = quantile(.epred, probs = 0.025), # 2.5% quantile (lower bound)
    .epred_high = quantile(.epred, probs = 0.975) # 97.5% quantile (upper bound)
  )
```




